---
title: "Control function flow"
format: html
---

# Instructions

This week, you learned how to work with lists, iterate with purrr, and control function flow with conditional statements. In this milestone, you'll use what you've learned to write a new function and then iterate over a vector of inputs to that function.

To learn more about the variables contained in the data set, refer to the data dictionary: https://rsacdn.link/milestones/internal/pinr/census/assets/census_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false
source("census_02_conditionals.R")
# Load your packages here

```

## Milestone 2

In this milestone, you'll write a function to summarize variables for each geographic division. Then, you'll iterate over a vector of variables to produce many summary tables at once.

## Recreation

### Part 1 - Summarize outcomes for each division

First, in the script `census_02_conditionals.R`, create a function named `summarize_divisions()`. `summarize_divisions()` should take two arguments:

* `data`: A processed ACS data set, like the result of `read_year("data/acs-2019.csv")`.
* `var`: A string containing the name of a variable, like `total_pop`.

Then, `summarize_divisions()` should create a summary table displaying the total `var` value for each division in `data`, like this:

```{r}
#| label: recreate-this
#| echo: false
solution <- readr::read_csv("solutions/solution_02_pt1.csv")
solution
```

We’ll soon learn some neat tricks for working with variables in a data set. Until then, we already know how to work with column names as strings. One method is to use `pivot_longer()` in concert with `pivot_wider()`, as shown in the following hint. (Note: You will need to add a number of other data wrangling steps to this pipeline.)

```{r}
#| label: get-started
#| eval: false
data |> 
  # select `division` and all numeric variables; remove `year`
  select(division, where(is.numeric), -year) |>
  # pivot the data into a longer format
  pivot_longer(-division) |>
  # filter to rows for your chose variable, `var`
  filter(name == var) |> 
  # add more data wrangling steps below...
  ____
```

*Hint 1:* Don't forget to group and summarize the data and arrange rows in the right order to match the solution.

*Hint 2:* You will need a `pivot_wider()` step at the end of your data wrangling pipeline

When you're done, un-comment and run the following code to see if your function works as expected.

```{r}
#| label: compare-1
#| message: false

# acs_2019 <- read_year("data/acs-2019.csv")
# result <- summarize_divisions(acs_2019, var = "total_pop")
# waldo::compare(
#   result, 
#   solution, 
#   tolerance = 1e-4, 
#   ignore_attr = c("class", "groups")
# )
```

### Part 2 - Iterate

Next, use `purrr::map()` and `summarize_divisions()`, setting `data` to `acs_2019` to create a summary table for each outcome variable in the following vector:

```{r}
#| label: outcome-vec
#| eval: false
c("employer_health_insurance", "medicare", "medicaid")
```

Write your code below.

```{r}
#| label: recreation-2

```

## Extension

Go even further. Make plans to apply techniques from week 2 or try something completely new. If you're looking for inspiration, consider the following suggestions:

* Some variables might better be summarized by taking the mean or median. Add an argument to `summarize_divisions()` to allow the choice of some summary function other than `sum()`. 
* Write a complementary function called `summarize_strata()` to work with data from supplemental data files like "ambulatory_difficulty.csv", "medicaid.csv", and others containing data divided by demographic strata.
* Find a new package to work with the data or to add something to a visualization, incorporating it into a new function.
* Write another function (or functions) to further process or explore the ACS data sets.
* Introduce `map()` to improve some aspect of your working with data outside Academy.
